---
title: "R interface for MLeap"
output:
  github_document:
    fig_width: 9
    fig_height: 5
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = TRUE
)

Sys.setenv("JAVA_HOME" = "/Library/Java/JavaVirtualMachines/jdk-18.0.1.1.jdk/Contents/Home")
rJava::.jinit() 
library(mleap)


Sys.setenv("JAVA_HOME" = "/usr/local/java")

```

[![Travis build status](https://travis-ci.org/rstudio/mleap.svg?branch=master)](https://travis-ci.org/rstudio/mleap) [![Coverage status](https://codecov.io/gh/rstudio/mleap/branch/master/graph/badge.svg)](https://codecov.io/github/rstudio/mleap?branch=master) [![CRAN status](https://www.r-pkg.org/badges/version/mleap)](https://cran.r-project.org/package=mleap)

**mleap** is a [sparklyr](http://spark.rstudio.com/) extension that provides an interface to [MLeap](https://github.com/combust/mleap), which allows us to take Spark pipelines to production.

## Getting started

**mleap** can be installed from CRAN via

```{r eval = FALSE}
install.packages("mleap")
```

or, for the latest development version from GitHub, using

```{r eval = FALSE}
devtools::install_github("rstudio/mleap")
```

Once mleap has been installed, we can install the external dependencies using

```{r eval = FALSE}
library(mleap)
install_maven()
# Alternatively, if you already have Maven installed, you can 
#  set options(maven.home = "path/to/maven")
install_mleap()

```


We can now export Spark ML pipelines from sparklyr.

```{r, message = FALSE}
library(sparklyr)
sc <- spark_connect("local", version = "3.2")

mtcars_tbl <- copy_to(sc, mtcars)

# Create a pipeline and fit it
pipeline <- ml_pipeline(sc) %>%
  ft_binarizer("hp", "big_hp", threshold = 100) %>%
  ft_vector_assembler(c("big_hp", "wt", "qsec"), "features") %>%
  ml_gbt_regressor(label_col = "mpg")

pipeline_model <- ml_fit(pipeline, mtcars_tbl)

# Export model
model_path <- file.path("mtcars_model.zip")

preds <- ml_transform(pipeline_model, mtcars_tbl)

ml_write_bundle(
  x = pipeline_model,  
  sample_input = mtcars_tbl, 
  path = model_path, 
  overwrite = TRUE
  )



# Disconnect from Spark
spark_disconnect(sc)
```

At this point, we can share `mtcars_model.zip` with our deployment/implementation engineers, and they would be able to embed the model in another application. See the [MLeap docs](http://mleap-docs.combust.ml/) for details.

We also provide R functions for testing that the saved models behave as expected. Here we load the previously saved model:

```{r}
model <- mleap_load_bundle(model_path)
model
```

We can retrieve the schema associated with the model:

```{r}
mleap_model_schema(model)
```

Then, we create a new data frame to be scored, and make predictions using our model:

```{r}
newdata <- tibble::tribble(
  ~qsec, ~hp, ~wt,
  16.2,  101, 2.68,
  18.1,  99,  3.08
)

# Transform the data frame
transformed_df <- mleap_transform(model, newdata)
dplyr::glimpse(transformed_df)
```

